\section{The Algebraic and Order Limit Theorems}

\begin{exercise}
  Let $x_{n} \geq 0$ for all $n \in \mathbf{N}$.
  \enum{
  \item If $\left(x_{n}\right) \rightarrow 0$, show that $\left(\sqrt{x_{n}}\right) \rightarrow 0$.
  \item If $\left(x_{n}\right) \rightarrow x$, show that $\left(\sqrt{x_{n}}\right) \rightarrow \sqrt{x}$.
  }
\end{exercise}

\begin{solution}
  \enum{
  \item
  Given some $\epsilon>0$, choose $N$ such that $x_n<\epsilon^2$ for all $n\geq N$. Then $\sqrt x_n < \epsilon$ for all $n\geq N$. 
  \item
  Fix $\epsilon>0$; now consider $\delta = \epsilon\sqrt x$. For some $N\in\mathbf{N}$, $|x_n - x|<\delta$ for all $n\geq N$, which implies $$|\sqrt x_n - \sqrt x| = \frac{|x_n-x|}{\sqrt x_n + \sqrt x} \leq \frac{|x_n-x|}{\sqrt x} < \frac{\delta}{\sqrt x}.$$ From this, we have $|\sqrt x_n -\sqrt x| < \epsilon$ for all $n \geq N$ as desired.
  }
\end{solution}

\begin{exercise}
  Using only Definition 2.2.3, prove that if $\left(x_{n}\right) \rightarrow 2$, then
  \enum{
  \item $\left(\frac{2 x_{n}-1}{3}\right) \rightarrow 1$;
  \item $\left(1 / x_{n}\right) \rightarrow 1 / 2$.
  }
  (For this exercise the Algebraic Limit Theorem is off-limits, so to speak.)
\end{exercise}

\begin{solution}
  \enum{
  \item 
  Let $\delta = \frac{3}{2}\epsilon$. Then 

  $$
  \begin{aligned}
  |x_n-2|<\delta
  &\implies \frac{2}{3}|x_n-2|<\epsilon \\
  &\implies \left| \frac{2x_n-4}{3} \right| < \epsilon\\
  &\implies \left| \frac{2x_n-1}{3} - 1 \right| < \epsilon \text{ for all } n\geq N \text{, as desired.}
  \end{aligned}
  $$

  \item 
  Note that
  $$\left| \frac{1}{x_n} - \frac{1}{2} \right| = \left| \frac{x_n-2}{2x_n} \right|.$$
  Choose $N$ large enough such that $x_n > 1$ and $|x_n-2|<\epsilon$ for all $n\geq N$. Then $$\left| \frac{x_n-2}{2x_n}\right|<\frac{|x_n-2|}{2}<\frac{\epsilon}{2}<\epsilon.$$
  }
\end{solution}

\begin{exercise}[Squeeze Theorem]
  Show that if $x_{n} \leq y_{n} \leq z_{n}$ for all $n \in \mathbf{N}$, and if $\lim x_{n}=\lim z_{n}=l$, then $\lim y_{n}=l$ as well.
\end{exercise}

\begin{solution}
  For any $\epsilon > 0$, 
  $$
  \begin{array}{l}
  |y_n - l|\leq |z_n -l|<\epsilon \text{ if } y_n-l\geq 0\\
  |y_n - l|\leq |x_n -l|<\epsilon \text{ otherwise}.
  \end{array}
  $$
\end{solution}

\begin{exercise}
  Let $\left(a_{n}\right) \rightarrow 0$, and use the Algebraic Limit Theorem to compute each of the following limits (assuming the fractions are always defined):
  \enum{
  \item $\lim \left(\frac{1+2 a_{n}}{1+3 a_{n}-4 a_{n}^{2}}\right)$
  \item $\lim \left(\frac{\left(a_{n}+2\right)^{2}-4}{a_{n}}\right)$
  \item $\lim \left(\frac{\frac{2}{a_{n}}+3}{\frac{1}{a_{n}}+5}\right)$
  }
\end{exercise}

\begin{solution}
  \enum{
  \item 
  $$
  \begin{aligned}
  \lim \left(\frac{1+2 a_{n}}{1+3 a_{n}-4 a_{n}^{2}}\right)
  &= \frac{\lim \left(1+2a_n\right)}{\lim(1+3a_n-4a_n^2)} \\
  &= \frac{\lim 1 + 2\lim a_n}{\lim 1 + 3\lim a_n - 4\lim a_n^2} \\
  &= 1
  \end{aligned}
  $$
  
  \item
  $$
  \begin{aligned}
  \lim \left(\frac{\left(a_{n}+2\right)^{2}-4}{a_{n}}\right)
  &= \lim \left( \frac{a_n^2+4a_n}{a_n} \right) \\
  &= \lim \left( a_n + 4 \right) \\
  &= 4
  \end{aligned}
  $$

  \item
  $$
  \begin{aligned}
  \lim \left(\frac{\frac{2}{a_{n}}+3}{\frac{1}{a_{n}}+5}\right) 
  &= \lim \left( \frac{2+3a_n}{1+5a_n} \right) \\
  &= \frac{2+3 \lim a_n}{1+5\lim a_n} \\
  &= 2
  \end{aligned}
  $$
  }
\end{solution}

\begin{exercise}
  Let $\left(x_{n}\right)$ and $\left(y_{n}\right)$ be given, and define $\left(z_{n}\right)$ to be the ``shuffled'' sequence $\left(x_{1}, y_{1}, x_{2}, y_{2}, x_{3}, y_{3}, \ldots, x_{n}, y_{n}, \ldots\right)$. Prove that $\left(z_{n}\right)$ is convergent if and only if $\left(x_{n}\right)$ and $\left(y_{n}\right)$ are both convergent with $\lim x_{n}=\lim y_{n}$.
\end{exercise}

\begin{solution}
  First we prove the forward direction. Let $\lim x_n = \lim y_n = l$, and fix $\epsilon>0$. Then there exists an $N_0\in\mathbf{N}$ such that $|x_n - l| <\epsilon$ and $|y_n-l|<\epsilon$ for all $n\geq N_0$. Setting $N=2N_0-1$, however, leaves us with the tail $(x_n, y_n, x_{n+1}, y_{n+1}, \ldots)$, which clearly satisfies $|z_n-l|<\epsilon$.
  \par
  For the backward direction, fix $\epsilon$. Since $|z_n - l| <\epsilon$ for all $n\geq N$, $|x_n - l| <\epsilon$ and $|y_n-l|<\epsilon$ for all $n \geq \frac{N}{2} + 1$.
\end{solution}

\begin{exercise}
  Consider the sequence given by $b_{n}=n-\sqrt{n^{2}+2 n}$. Taking $(1 / n) \rightarrow 0$ as given, and using both the Algebraic Limit Theorem and the result in Exercise 2.3.1, show $\lim b_{n}$ exists and find the value of the limit.
\end{exercise}

\begin{solution}
\BAD
$$
\begin{aligned}
b_n
&= n-\sqrt{n^2-2n} \\
&= \sqrt{n^2} - \sqrt{n^2-2n} \\
&= \frac{2n}{\sqrt{n^2}+\sqrt{n^2-2n}} \\
&= \frac{2}{1+\sqrt{1-\frac{2}{n}}}
\\
\lim b_n
&= \frac{2}{\lim{1+\sqrt{1-\frac{2}{n}}}} \\
&= \frac{2}{1+\lim\left(\sqrt{1-\frac{2}{n}}\right)} \\
&= \frac{2}{1+\sqrt{\lim \left( 1-\frac{2}{n} \right)}} \\
&= \frac{2}{1+1} \\
&= 1
\end{aligned}
$$
\end{solution}

\begin{exercise}
  Give an example of each of the following, or state that such a request is impossible by referencing the proper theorem(s):
  \enum{
  \item sequences $\left(x_{n}\right)$ and $\left(y_{n}\right)$, which both diverge, but whose sum $\left(x_{n}+y_{n}\right)$ converges;
  \item sequences $\left(x_{n}\right)$ and $\left(y_{n}\right)$, where $\left(x_{n}\right)$ converges, $\left(y_{n}\right)$ diverges, and $\left(x_{n}+y_{n}\right)$ converges;
  \item a convergent sequence $\left(b_{n}\right)$ with $b_{n} \neq 0$ for all $n$ such that $\left(1 / b_{n}\right)$ diverges;
  \item an unbounded sequence $\left(a_{n}\right)$ and a convergent sequence $\left(b_{n}\right)$ with $\left(a_{n}-b_{n}\right)$ bounded;
  \item two sequences $\left(a_{n}\right)$ and $\left(b_{n}\right)$, where $\left(a_{n} b_{n}\right)$ and $\left(a_{n}\right)$ converge but $\left(b_{n}\right)$ does not.
  }
\end{exercise}

\begin{solution}
  \enum{
  \item $a_n = (-1)^n$ and $b_n = (-1)^{n+1}$.
  \item This is impossible. By the Algebraic Limit Theorem, $\lim(x_n + y_n) - \lim(x_n) = \lim(y_n)$, and if $x_n + y_n$ and $x_n$ converge, then $y_n$ must have a limit as well.
  \item $b_n = 1/n$. Here, $1/b_n = n$, which diverges.
  \item There exist $M_0$ and $M_1$ such that $|b_n|<M_0$ and $|a_n-b_n|<M_1$. Therefore $|a_n-b_n|+|b_n|<M_0+M_1$, and by the triangle inequality, $$|a_n-b_n+b_n|<|a_n+b_n|+|b_n|<M_0+M_1$$ Thus $a_n$ is bounded by $M_0+M_1$, and an unbounded sequence that meets the criteria is impossible.
  \item $a_n=0$ and $b_n=(-1)^n$.
  }
\end{solution}

\begin{exercise}
  Let $\left(x_{n}\right) \rightarrow x$ and let $p(x)$ be a polynomial.
  \enum{
  \item Show $p\left(x_{n}\right) \rightarrow p(x)$.
  \item Find an example of a function $f(x)$ and a convergent sequence $\left(x_{n}\right) \rightarrow x$ where the sequence $f\left(x_{n}\right)$ converges, but not to $f(x)$.
  }
\end{exercise}

\begin{solution}
  \enum{
  \item
  Every polynomial has the form $p(x) = a_k x_n^k + \ldots + a_0$. 
  $$
  \begin{aligned}
  \lim p(x_n)
  &= \lim a_k x_n^k +\ldots +\lim a_0 \\
  &= a_k \lim x_n^k + \ldots + \lim a_0 \\
  &= a_k x^k + \ldots + a_0 \\
  &= p(x)
  \end{aligned}
  $$
  \item $x_n = 1/n$, and let $f(x)$ be a function where if $x\neq 0$, $f(x)=2$ and $f(x)=3$ otherwise. Then $(x_n)\to 0$, $f(0)=3$,  but $f(x_n)$ really converges to $2$.
  }
\end{solution}

\begin{exercise}
  \enum{
  \item Let $\left(a_{n}\right)$ be a bounded (not necessarily convergent) sequence, and assume $\lim b_{n}=0$. Show that $\lim \left(a_{n} b_{n}\right)=0$. Why are we not allowed to use the Algebraic Limit Theorem to prove this?
  \item Can we conclude anything about the convergence of $\left(a_{n} b_{n}\right)$ if we assume that $\left(b_{n}\right)$ converges to some nonzero limit $b$ ?
  \item Use (a) to prove Theorem 2.3.3, part (iii), for the case when $a=0$.
  }
\end{exercise}

\begin{solution}
  \enum{
  \item 
  We aren't allowed to use the Algebraic Limit Theorem because it assumes that $a_n$ and $b_n$ converge.

  Given that $|a_n|<M$ and $(b_n)\to 0$, then $|a_n b_n| < M|b_n| < M\epsilon$.

  \item 
  We can say that $a_n b_n$ is bounded, but we can't say anything about its convergence.
  \item 
  Convergent sequences are also bounded, so $b_n$ satisfies the conditions in part (a).
  }
\end{solution}

\begin{exercise}
  Consider the following list of conjectures. Provide a short proof for those that are true and a counterexample for any that are false.
  \enum{
  \item If $\lim \left(a_{n}-b_{n}\right)=0$, then $\lim a_{n}=\lim b_{n}$.
  \item If $\left(b_{n}\right) \rightarrow b$, then $\left|b_{n}\right| \rightarrow|b|$.
  \item If $\left(a_{n}\right) \rightarrow a$ and $\left(b_{n}-a_{n}\right) \rightarrow 0$, then $\left(b_{n}\right) \rightarrow a$.
  \item If $\left(a_{n}\right) \rightarrow 0$ and $\left|b_{n}-b\right| \leq a_{n}$ for all $n \in \mathbf{N}$, then $\left(b_{n}\right) \rightarrow b$.
  }
\end{exercise}

\begin{solution}
  \enum{
  \item This is necessarily true only if $a_n$ and $b_n$ are convergent, by the Algebraic Limit Theorem. Otherwise, consider $a_n = b_n = (-1)^n$.
  \item True, since $||b_n| - |b|| \leq |b_n - b| < \epsilon$.
  \item By the Algebraic Limit Theorem, $\lim b_n = \lim (b_n - a_n) + \lim a_n = a$.
  \item True, since $|b_n -b| \leq a_n < \epsilon$ for $n\geq N$. 
  }
\end{solution}

\begin{exercise}[Cesaro Means]
  \enum {
  \item Show that if $\left(x_{n}\right)$ is a convergent sequence, then the sequence given by the averages
    $$
    y_{n}=\frac{x_{1}+x_{2}+\cdots+x_{n}}{n}
    $$
    also converges to the same limit.
  \item Give an example to show that it is possible for the sequence $\left(y_{n}\right)$ of averages to converge even if $\left(x_{n}\right)$ does not.
  }
\end{exercise}

\begin{solution}
  \enum{
  \item 
  Fix some $\epsilon > 0$. First, choose $N$ such that $|x_n - x| < \epsilon/2$ for all $n > N$. Let $M$ be the sum of the first $N$ terms of $|x_n - x|$. Now choose some $X$ such that 
  $$
  \begin{aligned}
  \frac{2M}{\epsilon} - N < X
  &\implies M - N\epsilon < \frac{X\epsilon}{2} \\
  &\implies M + \frac{X\epsilon}{2} < (N+X)\epsilon \\
  &\implies |x_1-x|+\ldots+|x_{N+X}-x| < M + \frac{X\epsilon}{2} < (N+X)\epsilon \\
  &\implies |\frac{x_1 + \ldots + x_{N+X}}{N+X} - x|<\epsilon
  \end{aligned}
  $$
  Essentially, $X$ is the number of terms $|x_n - x| < \epsilon/2$ that we need to "drag" the mean back under $\epsilon$.

  Then it is easy to see that for all $n\geq N+X$, $|y_n - x| <\epsilon$.
  \item 
  $a_n = (-1)^n$, where the sequence of means converges to $0$.
  }
\end{solution}

\begin{exercise}
  A typical task in analysis is to decipher whether a property possessed by every term in a convergent sequence is necessarily inherited by the limit. Assume $\left(a_{n}\right) \rightarrow a$, and determine the validity of each claim. Try to produce a counterexample for any that are false.

  \enum{
  \item If every $a_{n}$ is an upper bound for a set $B$, then $a$ is also an upper bound for $B$.
  \item If every $a_{n}$ is in the complement of the interval $(0,1)$, then $a$ is also in the complement of $(0,1)$.
  \item If every $a_{n}$ is rational, then $a$ is rational.
  }
\end{exercise}

\begin{solution}
  \enum{
  \item
  Since $B$ has at least one upper bound, it is bounded and therefore has a supremum. Call this supremum $b$. If every $a_n$ is an upper bound for $B$, $a_n\geq b$ for all $n\in\mathbf{N}$. By the Order Limit Theorem, $\lim a_n = a\geq b$, so $a$ is an upper bound on $B$.
  \item
  This is false. Consider $a_n = -\frac{1}{n}$, which is always negative and therefore in the complement of $(0, 1)$ for all finite $n$, but tends to $0$.
  \item 
  This is false. Consider decimal approximations of $\sqrt 2$ of increasing length: they are all rational, but tend to an irrational number.
  }
\end{solution}

\begin{exercise}[Iterated Limits]
  Given a doubly indexed array $a_{m n}$ where $m, n \in \mathbf{N}$, what should $\lim _{m, n \rightarrow \infty} a_{m n}$ represent?
  \enum{
  \item Let $a_{m n}=m /(m+n)$ and compute the iterated limits
  $$
  \lim _{n \rightarrow \infty}\left(\lim _{m \rightarrow \infty} a_{m n}\right) \quad \text { and } \lim _{m \rightarrow \infty}\left(\lim _{n \rightarrow \infty} a_{m n}\right)
  $$
  Define $\lim _{m, n \rightarrow \infty} a_{m n}=a$ to mean that for all $\epsilon>0$ there exists an $N \in \mathbf{N}$ such that if both $m, n \geq N$, then $\left|a_{m n}-a\right|<\epsilon$.
  \item Let $a_{m n}=1 /(m+n)$. Does $\lim _{m, n \rightarrow \infty} a_{m n}$ exist in this case? Do the two iterated limits exist? How do these three values compare? Answer these same questions for $a_{m n}=m n /\left(m^{2}+n^{2}\right)$.
  \item Produce an example where $\lim _{m, n \rightarrow \infty} a_{m n}$ exists but where neither iterated limit can be computed.
  \item Assume $\lim _{m, n \rightarrow \infty} a_{m n}=a$, and assume that for each fixed $m \in \mathbf{N}$, $\lim _{n \rightarrow \infty}\left(a_{m n}\right) \rightarrow b_{m}$. Show $\lim _{m \rightarrow \infty} b_{m}=a$
  \item Prove that if $\lim _{m, n \rightarrow \infty} a_{m n}$ exists and the iterated limits both exist, then all three limits must be equal.
  }
\end{exercise}

\begin{solution}
  \enum{
  \item
  $$\lim_{n\to\infty} \left( \lim_{m\to\infty} a_{mn} \right) = 1$$
  $$\lim_{m\to\infty} \left( \lim_{n\to\infty} a_{mn} \right) = 0$$
  \item
  For $a_mn = 1/(m+n)$, all three limits are zero. For $a_mn = mn/(m^2+n^2)$, the combined limit does not exist because $a_mn = \frac{1}{2}$ whenever $M = N$, whereas the iterated limits are zero.
  \item 
  \REPLACE
  $$a_mn = \frac{(-1)^n}{m} + \frac{(-1)^m}{n}.$$
  \item 
  Fix $\epsilon > 0$. We are given that, for some $S_0$ and some $S_1$, 
  $$
  |a_{mn} - b_m| < \epsilon/2 \text{ for } m\in\mathbf{N}, n\geq S_0
  $$
  $$
  |a_{mn} - a|<\epsilon/2 \text{ for } m, n\geq S_1
  $$
  Now choose $N = \max(S_0, S_1)$. Then 
  $$
  |b_m-a|\leq|b_m-a_{mn}|+|a_{mn}-a|<\epsilon \text { for } m \geq N.
  $$
  \item 
  This was already shown in (d).
  }
\end{solution}